# Leave one out analysis in human

In primate brain data, human has four biological replicates and the NHPs only have three. Therefore, we make sure that the amount of correlation is not cause by higher biological replicate but because of species. However, our results suggest that different number will affect the correlation results. Therefore, we consider correlations that only existed in all four biological samples.

```{r warning=FALSE, message=FALSE, eval=FALSE}
library(TEKRABber)
library(twice)
library(dplyr)
```

We first load the metadata information and select sample ID within cluster 1 in human.

```{r eval=FALSE}
data("hmKZNFs337")

meta_combine <- metadata %>% inner_join(brain_meta, join_by(brain_region==region))

# prepare individual ID list
group_noHa <- meta_combine %>%
    filter(individual %in% c("hb", "hc", "hd")) %>%
    filter(cluster=="cluster1")

group_noHb <- meta_combine %>%
    filter(individual %in% c("ha", "hc", "hd")) %>%
    filter(cluster=="cluster1")

group_noHc <- meta_combine %>%
    filter(individual %in% c("ha", "hb", "hd")) %>%
    filter(cluster=="cluster1")

group_noHd <- meta_combine %>%
    filter(individual %in% c("ha", "hb", "hc")) %>%
    filter(cluster=="cluster1")

```

Then we convert the expression raw counts to TPM value. For genes, we select KRAB-ZNFs after the converting.

```{r eval=FALSE}
# genes convert to tpm
hmGene_temp <- hmGene[,c(-1)]
sample_counts <- colSums(hmGene_temp)

scaling_factor <- sample_counts / 1e6

hmGene_tpm <- hmGene_temp
hmGene_tpm <- t(t(hmGene_tpm)/ scaling_factor + 1) * 1e6
hmGene_tpm <- as.data.frame(hmGene_tpm)
rownames(hmGene_tpm) <- hmGene$geneID

# sebset only KRAB-ZNFs
kznfs_tpm <- hmGene_tpm %>%
    filter(rownames(.) %in% hmKZNFs337$ensembl_gene_id) %>%
    mutate(name = rownames(.)) %>%
    left_join(hmKZNFs337, join_by("name"=="ensembl_gene_id"))
rownames(kznfs_tpm) <- kznfs_tpm$external_gene_name
kznfs_tpm <- kznfs_tpm[,c(1:132)]


# tes convert to tpm
hmTE_temp <- hmTE[,-c(1,2,3)]
te_count <- colSums(hmTE_temp)
te_scale <- te_count / 1e6
hmTE_tpm <- hmTE_temp
hmTE_tpm <- t(t(hmTE_tpm)/ te_scale + 1) * 1e6
hmTE_tpm <- as.data.frame(hmTE_tpm)
rownames(hmTE_tpm) <- hmTE$name
```

Create a function called `corr_human` which filter the human ID for us and then calculate the correlations.

```{r eval=FALSE}
corr_human <- function(human_id){
    
    df_gene <- kznfs_tpm %>% select(human_id$Run)
    df_te <- hmTE_tpm %>% select(human_id$Run)
    
    df_corr <- corrOrthologTE(
        geneInput = df_gene,
        teInput = df_te,
        numCore = 5
    )
    
    # let us assume the cutoff will be absolute cofficient larger than 0.3 and 
    # p-adj less than 0.01
    df_corr_sig <- df_corr %>%
        filter(padj<0.01 & abs(coef)>=0.4) %>%
        mutate(pair = paste0(teName, ":", geneName))
    
    neg_count <- df_corr_sig %>% filter(coef<0) %>% nrow()
    pos_count <- df_corr_sig %>% filter(coef>0) %>% nrow()
    total_count <- nrow(df_corr_sig)
    count_kznf <- length(unique(df_corr_sig$geneName))
    count_te <- length(unique(df_corr_sig$teName))
    
    print(paste0("negative correlation: ", neg_count))
    print(paste0("positive correlation: ", pos_count))
    print(paste0("all correlation: ", total_count))
    print(paste0("unique KRAB-ZNFs: ", count_kznf))
    print(paste0("unique TEs: ", count_te))
    
    df_corr_sig
    
}
```

Execute `corr_human` in cluster 1 leave one sample out analysis:

```{r eval=FALSE}
c1_noHa <- corr_human(group_noHa)
c1_noHb <- corr_human(group_noHb)
c1_noHc <- corr_human(group_noHc)
c1_noHd <- corr_human(group_noHd)

```

```{r eval=FALSE}
# calculate all 4 biological replicates
c1_id <- meta_combine %>% 
    filter(cluster=="cluster1" & Organism=="Homo sapiens") %>% 
    select(Run) %>% unlist()

c1_all <- corrOrthologTE(
    geneInput = kznfs_tpm[colnames(kznfs_tpm) %in% c1_id],
    teInput = hmTE_tpm[colnames(hmTE_tpm) %in% c1_id],
    numCore = 5
)

c1_all_sig <- c1_all %>% 
    filter(padj<0.01 & abs(coef)>=0.4) %>%
    mutate(pair = paste0(teName, ":", geneName))
```

| Sample      | -/+/all                 | unique KRAB-ZNFs | unique TEs |
|-------------|-------------------------|------------------|------------|
| all 4 human | 26810 / 100987 / 127797 | 325              | 882        |
| no Ha       | 11050 / 54735 / 65785   | 303              | 849        |
| no Hb       | 21835 / 87725 / 109560  | 315              | 859        |
| no Hc       | 20132 / 75636 / 95768   | 318              | 861        |
| no Hd       | 22494 / 102193 / 124687 | 317              | 878        |

```{r eval=FALSE}
df_enrich <- read.csv("~/github/randomGeneSet/ChIP_Exo_Tables/kznfs_TEs_ChIP_exo_modified.csv")
df_enrich <- df_enrich %>%
    mutate(pair = paste0(teName, ":", geneName))

intersect_pair <- Reduce(intersect, list(c1_noHa$pair, c1_noHb$pair, c1_noHc$pair, c1_noHd$pair, df_enrich$pair)) #335: 241 positive correlation and 94 negative correlations

write.csv(intersect_pair, file="../tables/c1_confirm_correlation.csv", row.names=F)
```

```{r eval=FALSE}
# test for cluster 2
meta_combine <- metadata %>% inner_join(brain_meta, join_by(brain_region==region))

# prepare individual ID list
group_noHa_c2 <- meta_combine %>%
    filter(individual %in% c("hb", "hc", "hd")) %>%
    filter(cluster=="cluster2")

group_noHb_c2 <- meta_combine %>%
    filter(individual %in% c("ha", "hc", "hd")) %>%
    filter(cluster=="cluster2")

group_noHc_c2 <- meta_combine %>%
    filter(individual %in% c("ha", "hb", "hd")) %>%
    filter(cluster=="cluster2")

group_noHd_c2 <- meta_combine %>%
    filter(individual %in% c("ha", "hb", "hc")) %>%
    filter(cluster=="cluster2")


# calculate all 4 biological replicates
c2_id <- meta_combine %>% filter(cluster=="cluster2") %>% select(Run) %>% unlist()

c2_all <- corrOrthologTE(
    geneInput = kznfs_tpm[colnames(kznfs_tpm) %in% c2_id],
    teInput = hmTE_tpm[colnames(hmTE_tpm) %in% c2_id],
    numCore = 5
)

c2_all_sig <- c2_all %>% 
    filter(padj<0.01 & abs(coef)>=0.4) %>%
    mutate(pair = paste0(teName, ":", geneName))

c2_noHa <- corr_human(group_noHa_c2)
c2_noHb <- corr_human(group_noHb_c2)
c2_noHc <- corr_human(group_noHc_c2)
c2_noHd <- corr_human(group_noHd_c2)

```

| Sample      | -/+/all               | unique KRAB-ZNFs | unique TEs |
|-------------|-----------------------|------------------|------------|
| all 4 human | 11475 / 38295 / 49770 | 306              | 852        |
| no Ha       | 2717 / 10006 / 12723  | 253              | 752        |
| no Hb       | 9773 / 25666 / 35439  | 281              | 783        |
| no Hc       | 4807 / 16536 / 21343  | 282              | 767        |
| no Hd       | 9028 / 40769 / 49797  | 291              | 823        |

```{r eval=FALSE}
intersect_pair_c2 <- Reduce(intersect, list(c2_noHa$pair, c2_noHb$pair, c2_noHc$pair, c2_noHd$pair, df_enrich$pair)) #64: 35 positive correlation and 29 negative correlations

write.csv(intersect_pair_c2, file="../tables/c2_confirm_correlation.csv", row.names=F)
```
