# Optimize Network (Simplify)

In this script, we try to simplify the network in our manuscript.

The first one will be the 276 network comparing human to bonobo

## 276 network

```{r}
data("hg19rmsk_info")
pbd_obj <- readRDS("../data/pbd_obj.rds")
hmc1 <- pbd_obj$hmc1_corr
ppc1 <- pbd_obj$ppc1_corr
ppc1_neg <- ppc1 %>% filter(coef < 0)

intersect276 <- hmc1 %>% 
    filter(coef>0) %>%
    filter(pair %in% ppc1_neg$pair) # 276

hmc1_276 <- hmc1 %>% filter(pair %in% intersect276$pair) # all positive
ppc1_276 <- ppc1 %>% filter(pair %in% intersect276$pair) # all negative
```

We first based on ZNF112 and ZNF528 to understand the network

create a variable `te_family_count` for drawing the network

```{r}
y_kznf <- kznf_infer %>% filter(age=="young")
te_family_count <- hmc1_276 %>%
    mutate(kznf_age=ifelse(geneName %in% y_kznf$age, "y", "o")) %>%
    mutate(te_age=ifelse(teName %in% te_infer$NM, "y", "o")) %>%
    mutate(link_age=ifelse(kznf_age=="o" & te_age=="o", "o", "y")) %>%
    left_join(hg19rmsk_info[,c(1,2)], join_by(teName==gene_id)) %>%
    select(c(geneName, teName, family_id, link_age))
    
```

Break the network down

\(1\) TEs that only connect to ZNF528

\(2\) TEs that only connect the ZNF112

\(3\) TEs that connect to both ZNF528 and ZNF112

\(4\) the rest

```{r}
te_both <- te_family_count %>%
    filter(geneName %in% c("ZNF528", "ZNF112")) %>%
    filter(duplicated(teName)) #81

te_528 <- te_family_count %>%
    filter(geneName=="ZNF528") %>%
    filter(!teName %in% te_both$teName) #54

te_112 <- te_family_count %>%
    filter(geneName=="ZNF112") %>%
    filter(!teName %in% te_both$teName) #34
```

```{r}
te_both_count <- te_family_count %>%
    filter(geneName %in% c("ZNF528", "ZNF112")) %>%
    filter(teName %in% te_both$teName) %>%
    count(geneName, family_id)

```

Use cytoscape to draw our network

```{r}
# create a 276 network
library(netZooR)
library(RCy3)

te_family_count_modified <- te_family_count %>%
    mutate(family_id_modified = paste0(family_id, ":", geneName))

c1.276.kznf <- te_family_count_modified %>% select(geneName) %>% unique() #15
colnames(c1.276.kznf) <- c("id")
c1.276.te <- te_family_count_modified %>% 
    select(family_id_modified) %>% unique() #15
colnames(c1.276.te) <- c("id")
    
c1.276.node <- rbind(c1.276.kznf, c1.276.te)
c1.276.link <- te_family_count_modified[,c(1,5,3,4)]
colnames(c1.276.link) <- c("source", "target", "link_age", "n")

createNetworkFromDataFrames(c1.276.node, c1.276.link)
```
